---
title: "Ensemble Size - Preliminary Results"
author: "Rike"
date: "2023-07-01"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(dplyr)
options(dplyr.summarise.inform = FALSE)
library(here)
library(gt)
library(ggplot2)
DT <- `[`

nfcsts <- 100
plength <- 10
model_avail <- 0.8


ks <- c(4, 5, 6, 7)
loc <- "DE"
tt <- "Cases"

period_vec <- 2:5

source(here("R", "utils-ext.R"))

#read in forecast data
fcdat <- fread(here("data", "depldat.csv")) |>
  filter(location == "DE",
         target_type == "Cases") |>
  filter(period_cat %in% period_vec)

scoredat <- fread(here("data", "scores_period.csv")) |>
  filter(location == "DE", target_type == "Cases") |>
  select(period_cat, interval_score)


periods <- fcdat |>
  select(forecast_date, period_cat) |>
  distinct()

#for plotting
#exclude period 5
rcpreds <- data.table::fread(here("recombn-ens", "recombn-preds.csv")) |>
  filter(period_cat < 5)

rcscores <- data.table::fread(here("recombn-ens", "recombn-scores.csv"))|>
  mutate(k = as.factor(k))

fc_dates <- sort(unique(rcpreds$forecast_date))
fc_dates <- c(fc_dates, fc_dates[length(fc_dates)] + 7)
splitter <- c(1,10,10,10,10) |> cumsum()

date_list <- lapply(1:3, function(i) 
  c(fc_dates[splitter[i]], 
    fc_dates[(splitter[i+1])]))
  
#average size of ensemble over time
enssize <- fread(here("data", "depldat.csv")) |>
  filter(location == "DE", target_type == "Cases") |>
  select(model, forecast_date, location, target_type, period_cat) |>
  distinct() |>
  group_by(forecast_date) |>
  summarise(nmods = n()) |>
  left_join(periods, by = "forecast_date") |>
  group_by(period_cat) |>
  summarise(mean_avail = mean(nmods))
```

This document shows some preliminary scoring results for ensembles of size $k\in\{4,5,6,7\}$ that were recombined from all models available in the European Forecast Hub between `r fcdat$forecast_date |> min()` and `r fcdat$forecast_date |> max()` for \textbf{Cases} forecasts in \textbf{Germany}. 

## Score Table recombined ensembles

```{r, echo = FALSE}
rcscores |> 
  group_by(k, period_cat) |>
  summarise(meanscore = mean(ivscores),
            sdscore = sd(ivscores)) |>
  arrange(period_cat, k) |>
  select(period_cat, k, meanscore, sdscore) |>
  left_join(scoredat, by = "period_cat") |>
  mutate(rel_score = meanscore/interval_score,
         interval_score = NULL) |>
  left_join(enssize, by = "period_cat") |>
  ungroup() |>
  mutate(period_cat = paste0("Period ", period_cat, ", mean available models = ", mean_avail)) |>
  mutate(k = paste0("k = ", k)) |>
  select(period_cat, k, meanscore, rel_score, sdscore) |>
  gt(rowname_col = "k", groupname_col = "period_cat") |>
  cols_label(
      meanscore = "Average WIS",
      rel_score = "rel. to Ens.",
      sdscore = "Std. Dev. WIS"
    ) |>
  fmt_number(
    columns = c("meanscore", "sdscore"),
    decimals = 1,
    use_seps = FALSE
  )|>
  fmt_number(
    columns = c("rel_score"),
    decimals = 2,
    use_seps = FALSE
  )
```
## Trajectories of recombined ensembles


```{r, echo = FALSE, warning=FALSE, fig.dim = c(8, 12)}
plotdat <- rcpreds |>
  filter(quantile == 0.5, horizon == 2)

plotdat2 <- fread(here("data", "forecasts.csv")) |>
  filter(model == "EuroCOVIDhub-ensemble") |>
  filter(location == "DE", target_type == "Cases", 
         quantile == 0.5, horizon == 2) |>
  mutate(idens = 1) |>
  filter(forecast_date <= max(plotdat$forecast_date),
         forecast_date >= min(plotdat$forecast_date))

ggplot(aes(x = target_end_date, y = prediction, group = idens), data = plotdat) +
  geom_line(alpha = 0.1, color = "firebrick4") +
  ggplot2::geom_point(aes(x = target_end_date, y = true_value), data = plotdat, shape = 18) +
  geom_line(aes(x = target_end_date, y = value), data = plotdat2, color = "steelblue4", lwd = 1.005) +
  geom_line(aes(x = target_end_date, y = true_value), data = plotdat) +
  facet_wrap(~k, nrow = 4) +
  lapply(seq_along(date_list), function(dt) {
    list(annotate("rect", xmin = as.IDate(date_list[[dt]][1]), 
                  xmax = as.IDate(date_list[[dt]][2]), ymin = 0, ymax = Inf, 
                  alpha = .08 * (dt%%2 + 1)))
  }) +
  theme_masterthesis()
```


## Some boxplotty boxplots

Scores of recombined ensembles, relative to Hub ensemble. 

Each data point for boxplot is the sum of all $\textrm{WIS}$ scores^[yes, I know that's kinda like saying "Chai tea", but what you gonna do...] for the recombined Ensemble $Ens^R$ $i$ of size $k$ from period $p$, divided by the sum of all $\textrm{WIS}$ scores for the Hub Ensemble $HubEns$ for period $p$, i.e.
$$
\textrm{relWIS}(Ens^R_{k,p,i}) = \frac{\textrm{WIS}\left(Ens^R_{k,p,i}\right)}{\textrm{WIS}\left(HubEns_{p}\right)}
$$
This gives $100$ data points for each combination of period $p$ and size of recombination ensemble $k$, of which we can make boxplots: 

```{r pressure, echo=FALSE, warning=FALSE}
enssize_vec <- enssize |>
  filter(period_cat %in% period_vec) |>
  arrange(period_cat) |>
  select(mean_avail) |>
  pull()

names(enssize_vec) <- period_vec

hospital_labeller <- function(variable,value){
  return(paste0("avg. number of models in Hub-Ens: ", enssize_vec[value-1]))
}

ggplot(aes(x = k, y = rel_ivscores, fill = k), data = rcscores) +
  geom_boxplot() +
  geom_hline(yintercept = 1, linetype = "dashed") +
  facet_wrap(~period_cat, scales = "free",
             labeller = hospital_labeller) +
  theme_masterthesis() +
  ylab("WIS, relative to Hub ensemble") +
  xlab("") +
  scale_fill_brewer(palette = "Set2")
```

