---
title: "Ensemble model diversity"
output:
  github_document: default
---
### Identifying model types within varying ensemble compositions

#### Aim

- Comparing performance of ensembles by the methodology of component models
- Specifically comparing ensembles that combine across heterogeneous model types (among mechanistic, semi-mechanistic, or statistical), or whether the ensemble is of a single homogeneous model type 

```{r set-up, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(knitr)
library(here)
library(arrow)
library(purrr)
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(forcats)
theme_set(theme_bw())
plot_cols <- c("Mechanistic" = "#e08214", 
               "Semi-mechanistic" = "#c51b7d", 
                "Statistical" = "#4393c3",
                "Heterogeneous" = "grey50",
                "Homogeneous" = "#1b7837")
```

```{r get-targets}
# Get targets  ---------------------------------------------------
# -- keep only deaths as we only classified models forecasting deaths
source(here("ensvssize", "specs.R"))
targets <- crossing("loc" = enscomb_specs$loctargets,
                    "k" = enscomb_specs$ks) |>
  mutate(target = paste(loc, k, sep = "_k")) |>
  filter(grepl("Deaths", target)) |>
  pull(target)

# get model classifications
model_class <- read_csv("https://raw.githubusercontent.com/epiforecasts/eval-by-method/main/data/model-classification.csv") |>
  select(model, classification)
```
#### Methods

- `r length(unique(model_class$model))` Euro Hub models forecasting deaths, classified by 3 coders
- All available ensemble recombinations for 5 countries 
- Ensemble method is the median
- Performance metric is scaled relative skill against Hub median ensemble

#### Results

```{r get-model-ensemble-combinations}
# Ensemble combinations ---------------------------------------------------
# get ensemble component combinations
models <- map_dfr(.x = targets,
                   ~ read_parquet(here("enscomb-data",
                                       paste0("enscomb_",
                                              .x, ".parquet"))) |>
                     mutate(target = .x))

# Join to model classification
# - restrict to deaths as we only classified models with death forecasts
models <- models |>
  left_join(model_class, by = "model")

# count model classifications in each ensemble
ensemble_mix <- models |>
  group_by(ensid, target, k, classification) |>
  count() |>
  mutate(homog = k==n,
         classification = ifelse(!homog, "Heterogeneous", 
                                 classification)) |>
  select(-n) |>
  distinct()

# Get scores  ---------------------------------------------------
ensemble_method <- "hubreplica"
scores <- map_dfr(.x = targets,
                  ~ read_parquet(here("enscomb-data",
                                      paste0(ensemble_method, "-pwscores"),
                                      paste0("ens_comb_pwscores",
                                             .x, ".parquet"))) |>
                    mutate(target = .x))

scores <- scores |>
  filter(compare_against == "median-hubreplica" &
           model != "median-hubreplica") |>
  mutate(ensid = str_remove_all(model, "median_ensemble")) |>
  select(target, ensid, horizon, scaled_rel_skill)
```

```{r join-scores}
# Join scores and model type ----------------------------------------------
ensemble_scores <- left_join(scores, ensemble_mix,
                             by = c("ensid", "target")) |>
  filter(!is.na(classification))

######Rike's part - begin
# Count number of ensembles by homogeneous/heterogeneous type
num_ensembles <- ensemble_scores |> 
  select(target, homog, ensid, k) |> 
  distinct() |> 
  group_by(target, homog, k) |> 
  summarise(N = n()) |>
  mutate(location = substr(target, 1, 2)) |>
  mutate(target = substr(target, 3, 7)) |>
  mutate(homog = ifelse(homog, "Homogeneous", "Heterogeneous")) |>
  filter(k <= 5) |>
  mutate(k = paste0("k = ", k)) 
######Rike's part - end

# restrict to k < possible single-type ensembles
k_max <- filter(ensemble_scores, homog) |>
  group_by(target, horizon) |>
  summarise(k = max(k))

ensemble_scores <- filter(ensemble_scores,
                          target %in% k_max$target) |>
  # clean variables
  mutate(location = str_remove_all(target, "Deaths_k[:digit:]"),
         location = fct_infreq(location),
         horizon = ordered(horizon, levels = c(1,2),
                           labels = c("1 week", "2 week")),
         classification = fct_infreq(classification),
         homog = factor(homog,
                        levels = c(FALSE, TRUE),
                        labels = c("Heterogeneous", "Homogeneous")))
```

##### Available models

Unique models used in ensemble combinations, by method classification and target:

```{r plot-count, fig.dim=c(4,4)}
# Descriptive -------------------------------------------
# Count by classification
models |>
  separate_wider_delim(target, delim = "Deaths_k", names = c("loc", "k_")) |>
  distinct(loc, model, classification) |>
  mutate(loc = fct_infreq(loc)) |>
  group_by(loc, classification) |>
  count() |>
  pivot_wider(names_from = loc, values_from = n) |> 
  kable()
```


##### Ensemble scores by model classification

Distribution of all ensemble scores by model classification, across all (five) location targets:

```{r plot-by-type, fig.wid=8, fig.height=4}
# Boxplot ensemble scores by model classification ---------------
ensemble_scores |>
  ggplot(aes(y = classification)) +
  geom_boxplot(aes(x = scaled_rel_skill), varwidth = TRUE, outlier.alpha = 0.5) + 
  geom_vline(aes(xintercept = 1), lty = 2) +
  labs(y = NULL, x = "Scaled relative skill",
       subtitle = "Ensemble scores by model composition",
       caption = "Boxplot width scaled to the count of available ensembles") +
  facet_wrap(~horizon)
```

##### Ensemble scores by number of components

Plotting the mean and 90% range of performance of ensembles with increasing numbers of models, in comparison to the Hub ensemble (1).

- Plotting homogeneous ensembles by their individual model type:

```{r plot-summary-method}
# Summarise scores --------------------------------------------------
# code adapted from: plot_allrecresults.R
summary_scores_k_class <- ensemble_scores |>
  group_by(location, k, horizon, classification) |>
  summarise(
    n = n(),
    meanrelskill = mean(scaled_rel_skill),
    q05relskill = quantile(scaled_rel_skill, 0.05),
    q95relskill = quantile(scaled_rel_skill, 0.95))
# Plot side-by-side ---------------------------------------------------
summary_scores_k_class |>
  ggplot(aes(x = k, col = classification)) +
  geom_point(aes(y = meanrelskill),
             alpha = 0.8,
             position = position_dodge(width=0.5)) +
  geom_linerange(aes(ymin = q05relskill,
                     ymax = q95relskill),
                 alpha = 0.6,
                 linewidth = 2,
                 position = position_dodge(width=0.5)) +
  scale_x_continuous(breaks = seq(1:max(k_max$k))) +
  geom_hline(aes(yintercept = 1),
             linetype = 2, alpha = 0.5) +
  scale_color_manual(values = plot_cols) +
  labs(subtitle = "European ensemble forecasts, deaths",
       col = "Ensemble composition",
       x = NULL, y = "Scaled relative skill") +
  facet_grid(rows = vars(location),
             cols = vars(horizon),
             scales = "free_y") +
  theme(legend.position = "bottom",
        strip.background = element_rect(fill = NA, colour = NA))
```

- Plotting all homogeneous ensembles against all heterogeneous ensembles:

```{r plot-summary}
# Summarise scores --------------------------------------------------
# code adapted from: plot_allrecresults.R
summary_scores_k_homog <- ensemble_scores |>
  group_by(location, k, horizon, homog) |>
  summarise(
    n = n(),
    meanrelskill = mean(scaled_rel_skill),
    q05relskill = quantile(scaled_rel_skill, 0.05),
    q95relskill = quantile(scaled_rel_skill, 0.95))
# Plot side-by-side ---------------------------------------------------
summary_scores_k_homog |>
  ggplot(aes(x = k, col = homog)) +
  # geoms
  geom_point(aes(y = meanrelskill),
             alpha = 0.8,
             position = position_dodge(width=0.5)) +
  geom_linerange(aes(ymin = q05relskill,
                     ymax = q95relskill),
                 alpha = 0.6,
                 linewidth = 2,
                 position = position_dodge(width=0.5)) +
  # formatting
  scale_x_continuous(breaks = seq(1:max(k_max$k))) +
  geom_hline(aes(yintercept = 1),
             linetype = 2, alpha = 0.5) +
  scale_color_manual(values = plot_cols) +
  labs(subtitle = "European ensemble forecasts, deaths",
       col = "Ensemble composition",
       x = NULL, y = "Scaled relative skill") +
  facet_grid(rows = vars(location),
             cols = vars(horizon),
             scales = "free_y") +
  theme(legend.position = "bottom",
        strip.background = element_rect(fill = NA, colour = NA))
```



```{r num-ensembles}
###Rike's block
#Plot number of ensembles
ggplot(num_ensembles, aes(x = N, y = homog, fill = homog)) +
  geom_bar(stat = "identity") +
  facet_grid(location~ k, scales = "free") +
  scale_fill_manual(values = plot_cols) +
  #scale_fill_met_d("Hokusai2") +
  geom_text(aes(label = N), hjust = 0.6)  +
  guides(fill="none") +
  labs(subtitle = "Number of ensembles by type and k, deaths", 
       y = "")
```


```{r}
#### Model performance by type
```

